\textsc{RILS}-\textsc{ROLS} : Robust Symbolic Regression via Iterated Local Search and Ordinary Least Squares

### 2022-11-22 paper_resources/doc/paper.tex
Finished up to Algorithm 4.

\fxnote{Dati neku smislenu procjenu kompleksnosti algoritma, ako to nije pretesko. }
\subsubsection{Fitness function}\label{sec:fitness}

@@

This is followed by constant normalization during which coefficients that enter multiplication, division, addition or subtraction are set to 1, while those entering the power function are rounded to integer, with exception of square root, which is kept intact. For example, for expression $3.3\cdot(x+45.1\cdot y^{3.2})\cdot 81\cdot x/\sqrt{y}$ the normalized version is $1\cdot (x+1\cdot y^3)\cdot 1\cdot x/\sqrt{y}$. The reason for performing normalization is reducing the search space of possible perturbations thus keeping them within a reasonable size. This reduction is reasonable, since normalization preserves the essential functional form. (Note that coefficients get tuned later during the OLS phase and local search.)

After performing the normalization process, the expression is simplified -- getting compact expression is more likely after normalization than before. The previous expression will take form $(x+y^3)\cdot x/\sqrt{y}$. The purpose of simplification is removing unnecessary coefficients, but in general it can also be non-trivial symbolic simplification.

Perturbations are generated by making simple changes on the per-node level of $s$ expression tree.
Depending on the structure of the expression tree (note that expression does not have to have unique tree representation), the set of subtrees of the previous expression $(x+y^3)\cdot x/\sqrt{y}$ might be $\{(x+y^3)\cdot x/\sqrt{y}, (x+y^3), x/\sqrt{y}, x, y^3, \sqrt{y}, y\}$. %This set of subtrees is obtained if left subtree of the whole expression is $(x+y^3)$ while the right one is $x/\sqrt{y}$.  
For each subtree $n$, the set of perturbations around $n$ is generated (lines \ref{line:forSSStart}-\ref{line:forSSEnd} in Algorithm~\ref{alg:pert}).

>>>>

%\fxnote{Dati neku smislenu procjenu kompleksnosti algoritma, ako to nije pretesko. Aca: zasad bih ovo izbegao, nezgodno je. Ako budu trazili recenzenti, uradicemo.}

\subsubsection{Fitness function}\label{sec:fitness}

@@

This is followed by constant normalization during which coefficients that enter multiplication, division, addition or subtraction are set to 1, while those entering the power function are rounded to integer, with exception of square root, which is kept intact. For example, for expression $3.3\cdot(x+45.1\cdot y^{3.2})\cdot 81\cdot x/\sqrt{y}$ the normalized version is $1\cdot (x+1\cdot y^3)\cdot 1\cdot x/\sqrt{y}$. The reason for performing normalization is reducing the search space of possible perturbations. This reduction is reasonable, since normalization preserves the essential functional form. (Note that coefficients get tuned later: the linear coefficient during the OLS phase, and the remaining during local search.)

After performing the normalization process, the expression is simplified -- getting compact expression is more likely after normalization than before. The previous expression will take form $(x+y^3)\cdot x/\sqrt{y}$. In this particular case, the simplification will usually only remove unnecessary coefficients, but in general it can also perform some non-trivial symbolic simplification.

Perturbations are generated by making simple changes on the per-node level of $s$ expression tree.
Depending on the structure of the expression tree (note that expression does not have to have unique tree representation), the set of subtrees of the previous expression $(x+y^3)\cdot x/\sqrt{y}$ might be $\{(x+y^3)\cdot x/\sqrt{y}, (x+y^3), x/\sqrt{y}, x, y^3, \sqrt{y}, y\}$. %This set of subtrees is obtained if left subtree of the whole expression is $(x+y^3)$ while the right one is $x/\sqrt{y}$.  
Further, the set of perturbations around each subtree $n$ is generated (lines \ref{line:forSSStart}-\ref{line:forSSEnd} in Algorithm~\ref{alg:pert}).
